---
title: 'Radar Chart'
author: "Zehui Bai"
date: '`r format(Sys.time())`'
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
  pdf_document:
    toc: yes
fontsize: 10pt
editor_options:
  chunk_output_type: console
colorlinks: yes
---

```{r setup, include=FALSE, echo = FALSE,message = FALSE, error = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# <!-- ---------------------------------------------------------------------- -->
# <!--                    1. load the required packages                       -->
# <!-- ---------------------------------------------------------------------- --> 

## if(!require(psych)){install.packages("psych")}
# devtools::install_github("ricardo-bion/ggradar")

packages<-c("tidyverse", "knitr", "kableExtra", "papeR", "fmsb", "ggradar", "ggpubr")
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}
ipak(packages)
# <!-- ---------------------------------------------------------------------- --> 


# <!-- ---------------------------------------------------------------------- -->
# <!--                        2. Basic system settings                        -->
# <!-- ---------------------------------------------------------------------- -->
setwd(dirname(rstudioapi::getSourceEditorContext()$path))
getwd()
Sys.setlocale("LC_ALL","English")

## convert backslash to forward slash in R
# gsub('"', "", gsub("\\\\", "/", readClipboard()))

### get the path
# rstudioapi::getSourceEditorContext()$path
# dirname(rstudioapi::getSourceEditorContext()$path)

### set working directory
# getwd()
# setwd("c:/Users/zbai/Desktop")
# Sys.setlocale("LC_ALL","English")

### get the R Version
# paste(R.Version()[c("major", "minor")], collapse = ".")

### convert backslash to forward slash 
# scan("clipboard",what="string")
# gsub('"', "", gsub("\\\\", "/", readClipboard()))
# <!-- ---------------------------------------------------------------------- --> 
 
# <!-- ---------------------------------------------------------------------- -->
# <!--     3. Load the SASmarkdown package if the SAS output is required      -->
# <!-- ---------------------------------------------------------------------- -->
# library(SASmarkdown)
# ### Set SAS output
# ### Reset engine to R
# saspath <- "C:/SASHome/SASFoundation/9.4/sas.exe"
# sasopts <- "-nosplash -linesize 75"
# knitr::opts_chunk$set(engine="sashtml", engine.path=saspath,
#         engine.opts=sasopts, comment=NA)
# 
# # run these commands to convince yourself that
# # within this knitr session the engine changed.
# knitr::opts_chunk$get()$engine
# knitr::opts_chunk$get()$engine.path
# knitr::opts_chunk$get()$engine.opts
# <!-- ---------------------------------------------------------------------- -->



# <!-- ---------------------------------------------------------------------- -->
# <!--                         4. Import the datasets                         -->
# <!-- ---------------------------------------------------------------------- -->
### Import csv data
# pfad <- "~/Desktop/SASUniversityEdition/myfolders/Daten"
# mydata1 <- read.csv(file.path(pfad, "yourcsv_data.csv"), 
#                     sep=";", 
#                     header=TRUE)   

### Import xlsx data
# library(readxl)
# mydata2 <- read_excel("C:/Users/zbai/Documents/GitHub/R-Projects/SAS/Yimeng/results-text.xlsx")

### Import sas data
# library(sas7bdat)
# mydata3 <- read.sas7bdat("~/Desktop/SASUniversityEdition/myfolders/Daten/uis.sas7bdat")

### Import from copyboard
# copdat <- read.delim("clipboard")
# Data_D01 <- copdat

# <!-- ---------------------------------------------------------------------- -->
# <!--                           5. Some Tools                                -->
# <!-- ---------------------------------------------------------------------- -->

## To check out vignettes for one specific package
# browseVignettes("ggplot2")


# <!-- ---------------------------------------------------------------------- -->
```



```{r mind map,echo = F,message = FALSE, error = FALSE, warning = FALSE}
## Convert to mind map text, markdown outline, R script, and HTML widget ####
library(mindr)
# text -> widget
# input <- c("# Chapter 1", "## Section 1.1", "### Section 1.1.1", "## Section 1.2", "# Chapter 2")
# mm(from = input, root = "mindr")


input <- rstudioapi::getSourceEditorContext()$path 
## file.show(input) # Open the input file with the default program, if any
input_txt <- readLines(input, encoding = "UTF-8")
## Convert to mind map text, markdown outline, R script, and HTML widget ####
mm_output <- mm(input_txt, 
                output_type = c("widget"),
                root = "")
mm_output$widget
```



# Ref

* https://www.datanovia.com/en/blog/beautiful-radar-chart-in-r-using-fmsb-and-ggplot-packages/



# Radar plot use fmsb package

## Data

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
exam_scores <- data.frame(
    row.names = c("Student.1", "Student.2", "Student.3"),
      Biology = c(7.9, 3.9, 9.4),
      Physics = c(10, 20, 0),
        Maths = c(3.7, 11.5, 2.5),
        Sport = c(8.7, 20, 4),
      English = c(7.9, 7.2, 12.4),
    Geography = c(6.4, 10.5, 6.5),
          Art = c(2.4, 0.2, 9.8),
  Programming = c(0, 0, 20),
        Music = c(20, 20, 20)
)

exam_scores %>% kable(caption = "Demo Data", format = "html") %>%
  kable_styling(latex_options = "striped")

# Define the variable ranges: maximum and minimum
max_min <- data.frame(
  Biology = c(20, 0), Physics = c(20, 0), Maths = c(20, 0),
  Sport = c(20, 0), English = c(20, 0), Geography = c(20, 0),
  Art = c(20, 0), Programming = c(20, 0), Music = c(20, 0)
)
rownames(max_min) <- c("Max", "Min")

# Bind the variable ranges to the data
df <- rbind(max_min, exam_scores)
df %>% kable(caption = "Demo Data with variable ranges", format = "html") %>%
  kable_styling(latex_options = "striped")
```

## Radar plot 

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
student1_data <- df[c("Max", "Min", "Student.1"), ]
student1_data %>% radarchart()


# Helper function to produce a beautiful radar chart
create_beautiful_radarchart <- function(data, color = "#00AFBB", 
                                        vlabels = colnames(data), vlcex = 0.7,
                                        caxislabels = NULL, title = NULL, ...){
  radarchart(
    data, axistype = 1,
    # Customize the polygon
    pcol = color, pfcol = scales::alpha(color, 0.5), plwd = 2, plty = 1,
    # Customize the grid
    cglcol = "grey", cglty = 1, cglwd = 0.8,
    # Customize the axis
    axislabcol = "grey", 
    # Variable labels
    vlcex = vlcex, vlabels = vlabels,
    caxislabels = caxislabels, title = title, ...
  )
}

# Reduce plot margin using par()
op <- par(mar = c(1, 2, 2, 1))
create_beautiful_radarchart(student1_data, caxislabels = c(0, 5, 10, 15, 20))
par(op)


# Create radar charts for multiple individuals
# Reduce plot margin using par()
op <- par(mar = c(1, 2, 2, 2))
# Create the radar charts
create_beautiful_radarchart(
  data = df, caxislabels = c(0, 5, 10, 15, 20),
  color = c("#00AFBB", "#E7B800", "#FC4E07")
)
# Add an horizontal legend
legend(
  x = "bottom", legend = rownames(df[-c(1,2),]), horiz = TRUE,
  bty = "n", pch = 20 , col = c("#00AFBB", "#E7B800", "#FC4E07"),
  text.col = "black", cex = 1, pt.cex = 1.5
  )
par(op)


# Define colors and titles
colors <- c("#00AFBB", "#E7B800", "#FC4E07")
titles <- c("Student.1", "Student.2", "Student.3")

# Reduce plot margin using par()
# Split the screen in 3 parts
op <- par(mar = c(1, 1, 1, 1))
par(mfrow = c(1,3))

# Create the radar chart
for(i in 1:3){
  create_beautiful_radarchart(
    data = df[c(1, 2, i+2), ], caxislabels = c(0, 5, 10, 15, 20),
    color = colors[i], title = titles[i]
    )
}
par(op)
```


# ggplot radar chart using the ggradar R package


```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
# Put row names into  a column named group
df <- exam_scores %>% rownames_to_column("group")
df
```

## Basic radar plot

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
# Plotting student 1
ggradar(
  df[1, ], 
  values.radar = c("0", "10", "20"),
  grid.min = 0, grid.mid = 10, grid.max = 20
  )
```
  
## Customize radar charts

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
ggradar(
  df[1, ], 
  values.radar = c("0", "10", "20"),
  grid.min = 0, grid.mid = 10, grid.max = 20,
  # Polygons
  group.line.width = 1, 
  group.point.size = 3,
  group.colours = "#00AFBB",
  # Background and grid lines
  background.circle.colour = "white",
  gridline.mid.colour = "grey"
  )
```


## Radar chart with multiple individuals or groups 

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
ggradar(
  df, 
  values.radar = c("0", "10", "20"),
  grid.min = 0, grid.mid = 10, grid.max = 20,
  # Polygons
  group.line.width = 1, 
  group.point.size = 3,
  group.colours = c("#00AFBB", "#E7B800", "#FC4E07"),
  # Background and grid lines
  background.circle.colour = "white",
  gridline.mid.colour = "grey",
  legend.position = "bottom"
  )
```

# lolliplot/dotchart

A circular plot is difficult to read. An alternative to a radar chart is an ordered lolliplot or dotchart using ggpubr package

## Data preparation

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
df2 <- t(exam_scores) %>%
  as.data.frame() %>%
  rownames_to_column("Field")
df2

```

## Displaying one individual

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
ggdotchart(
  df2, x = "Field", y = "Student.1",
  add = "segments", sorting = "descending",
  ylab = "Exam Score", title = "Student 1"
  )
```

## Displaying two individuals

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
## Displaying two individuals
df3 <- df2 %>%
  select(Field, Student.1, Student.2) %>%
  pivot_longer(
    cols = c(Student.1, Student.2),
    names_to = "student",
    values_to = "value"
  )
ggdotchart(
  df3, x = "Field", y = "value", 
  group = "student", color = "student", palette = "jco",
  add = "segment", position = position_dodge(0.3),
  sorting = "descending"
  )
```

## Displaying multiple individuals

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
df4 <- df2 %>%
  select(Field, Student.1, Student.2, Student.3) %>%
  pivot_longer(
    cols = c(Student.1, Student.2, Student.3),
    names_to = "student",
    values_to = "value"
  )
head(df4)

ggdotchart(
  df4, x = "Field", y = "value", 
  group = "student", color = "student", palette = "jco",
  add = "segment", position = position_dodge(0.3),
  sorting = "descending", facet.by = "student",
  rotate = TRUE, legend = "none"
  )
```
 
 
