---
title: 'PMCF Survey China – RT Decalin Analysis'
author: "Zehui Bai"
date: 'Stand: `r format(Sys.time(), "%F %H:%M Uhr")`'
output:
  html_document:
    df_print: paged
    number_sections: no
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
  pdf_document:
    toc: yes
fontsize: 10pt
editor_options:
  chunk_output_type: console
colorlinks: yes
---



```{r setup, include=FALSE, echo = FALSE,message = FALSE, error = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# <!-- ---------------------------------------------------------------------- -->
# <!--                    1. load the required packages                       -->
# <!-- ---------------------------------------------------------------------- --> 
packages<-c("pwr","tidyverse", "knitr", "papeR")
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}
ipak(packages)
# <!-- ---------------------------------------------------------------------- --> 


# <!-- ---------------------------------------------------------------------- -->
# <!--                        2. Basic system settings                        -->
# <!-- ---------------------------------------------------------------------- -->
setwd(dirname(rstudioapi::getSourceEditorContext()$path))
getwd()
Sys.setlocale("LC_ALL","English")

### get the path
# rstudioapi::getSourceEditorContext()$path
# dirname(rstudioapi::getSourceEditorContext()$path)

### set working directory
# getwd()
# setwd("c:/Users/zbai/Desktop")
# Sys.setlocale("LC_ALL","English")

### get the R Version
# paste(R.Version()[c("major", "minor")], collapse = ".")

### convert backslash to forward slash 
# scan("clipboard",what="string")
# gsub('"', "", gsub("\\\\", "/", readClipboard()))
# <!-- ---------------------------------------------------------------------- --> 



# <!-- ---------------------------------------------------------------------- -->
# <!--     3. Load the SASmarkdown package if the SAS output is required      -->
# <!-- ---------------------------------------------------------------------- -->
# library(SASmarkdown)
# ### Set SAS output
# ### Reset engine to R
# saspath <- "C:/SASHome/SASFoundation/9.4/sas.exe"
# sasopts <- "-nosplash -linesize 75"
# knitr::opts_chunk$set(engine="sashtml", engine.path=saspath,
#         engine.opts=sasopts, comment=NA)
# 
# # run these commands to convince yourself that
# # within this knitr session the engine changed.
# knitr::opts_chunk$get()$engine
# knitr::opts_chunk$get()$engine.path
# knitr::opts_chunk$get()$engine.opts
# <!-- ---------------------------------------------------------------------- -->



# <!-- ---------------------------------------------------------------------- -->
# <!--                         4. Import the datasets                         -->
# <!-- ---------------------------------------------------------------------- -->
### Import csv data
# pfad <- "~/Desktop/SASUniversityEdition/myfolders/Daten"
# mydata1 <- read.csv(file.path(pfad, "yourcsv_data.csv"), 
#                     sep=";", 
#                     header=TRUE)   

### Import xlsx data
# library(readxl)
# mydata2 <- read_excel("C:/Users/zbai/Documents/GitHub/R-Projects/SAS/Yimeng/results-text.xlsx")

### Import sas data
# library(sas7bdat)
# mydata3 <- read.sas7bdat("~/Desktop/SASUniversityEdition/myfolders/Daten/uis.sas7bdat")
# <!-- ---------------------------------------------------------------------- -->


### Improt dataset from clipboard
# copdat <- read.delim("clipboard")
# Analysis_Domain_D01 <- copdat
# saveRDS(Analysis_Domain_D01, "RT_Decalin_Domains")
Analysis_Domain_D01 <- readRDS("RT_Decalin_Domains")
 

 
### Convert Q4: Dosage of Intraocular filling in ml   	
###         Q4: Dosage of lifting of subluxated lenses in ml	
###         Q4: Dosage of pressure out the subretinal fluid in ml 
### As numeric variables

Analysis_Domain_D02 <- Analysis_Domain_D01 %>%
  mutate(Q4ITM1 = case_when(Q4ITM1=="3-4"~"3.5",
                            Q4ITM1=="3-5"~"4",
                            Q4ITM1=="4"~"4",
                            Q4ITM1=="4-5"~"4.5",
                            Q4ITM1=="4-6"~"5",
                            Q4ITM1=="4.5"~"4.5",
                            Q4ITM1=="5"~"5",
                            Q4ITM1=="6"~"6",
                            Q4ITM1=="7"~"7"),
         Q4ITM2 = case_when(Q4ITM2=="0"~"0",
                            Q4ITM2=="2-3"~"2.5",
                            Q4ITM2=="3-4"~"3.5",
                            Q4ITM2=="3-5"~"4",
                            Q4ITM2=="4"~"4",
                            Q4ITM2=="4-5"~"4.5",
                            Q4ITM2=="5"~"5",
                            Q4ITM2=="7"~"7"),
         Q4ITM3 = case_when(Q4ITM3=="0"~"0",
                            Q4ITM3=="3"~"3", 
                            Q4ITM3=="3-4"~"3.5", 
                            Q4ITM3=="4"~"4",
                            Q4ITM3=="4-5"~"4.5",
                            Q4ITM3=="5"~"5",
                            Q4ITM3=="7"~"7"),
         Q4ITM1 = as.numeric(Q4ITM1),
         Q4ITM2 = as.numeric(Q4ITM2),
         Q4ITM3 = as.numeric(Q4ITM3)) 
```
 

# Descriptive analysis

## Demographic and Treatment Information

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}

xxx_replace_xxx_pcre <- function(string, patterns, replacements) {
   for (i in seq_along(patterns))
      string <- gsub(patterns[i], replacements[i], string, perl=TRUE)
   string
}
 

## cat(paste(colnames(read.delim("clipboard")), collapse=","), "\n")

  

### Select variables
Analysis_Domain_D03a <- Analysis_Domain_D02 %>%
  select(HOSCODE, Q1NCASE, 
         Q2SEXF, Q2SEXM,
         Q2AGEL, Q2AGEH, Q2AGECAT1, Q2AGECAT2, Q2AGECAT3,
         Q3DIS1, Q3DIS2, Q3DIS3, Q3DIS4OTH, Q3DIS4, Q3DIS5, Q3DIS6OTHa, Q3DIS6a,Q3DIS6OTHb, Q3DIS6b,
         Q4ITM1,Q4ITM2,Q4ITM3,Q4ITM4OTH,Q4ITM4) %>%
  mutate(Q2SEXF_N = Q2SEXF*Q1NCASE/100,
         Q2SEXM_N = Q2SEXM*Q1NCASE/100,
         Q2AGECAT1_N = Q2AGECAT1*Q1NCASE/100,
         Q2AGECAT2_N = Q2AGECAT2*Q1NCASE/100,
         Q2AGECAT3_N = Q2AGECAT3*Q1NCASE/100,
         Q3DIS1_N = Q3DIS1*Q1NCASE/100,
         Q3DIS2_N = Q3DIS2*Q1NCASE/100,
         Q3DIS3_N = Q3DIS3*Q1NCASE/100,
         Q3DIS4_N = Q3DIS4*Q1NCASE/100,
         Q3DIS5_N = Q3DIS5*Q1NCASE/100,
         Q3DIS6_N = (Q3DIS6a+Q3DIS6b)*Q1NCASE/100)

### Check the percentage
Analysis_Domain_D03a %>% filter(Q2SEXF+Q2SEXM!=100|
                                  Q2AGECAT1+Q2AGECAT2+Q2AGECAT3!=100|
                                  Q3DIS1+Q3DIS2+Q3DIS3+Q3DIS4+Q3DIS5+Q3DIS6a+Q3DIS6b!=100) %>%
  nrow()


### Hospitals participating in the survey 
N_hospital <- nrow(Analysis_Domain_D02)
N_hospital

### Performed operations using Perfluorocarbon RT DECALIN
N_case <- sum(Analysis_Domain_D02$Q1NCASE)
N_case

###  Gender
apply(Analysis_Domain_D03a %>% select(Q2AGECAT1_N,Q2AGECAT2_N,Q2AGECAT3_N), 2, sum)

###  Age (Range: 02-89 in years)
apply(Analysis_Domain_D03a %>% select(Q2SEXF_N, Q2SEXM_N), 2, sum)
min(Analysis_Domain_D03a$Q2AGEL)
max(Analysis_Domain_D03a$Q2AGEH)


### Disease requiring aid treatment with Perfluorocarbon RT DECALIN
apply(Analysis_Domain_D03a %>% select(Q3DIS1_N, Q3DIS2_N, Q3DIS3_N, Q3DIS4_N, Q3DIS5_N, Q3DIS6_N), 2, sum)
prop.table(apply(Analysis_Domain_D03a %>% select(Q3DIS1_N, Q3DIS2_N, Q3DIS3_N, Q3DIS4_N, Q3DIS5_N, Q3DIS6_N), 2, sum))

### Other retinal detachments
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}
firstup(names(table(Analysis_Domain_D03a$Q3DIS4OTH)))

### Other diseases
 
firstup(names(table(Analysis_Domain_D03a$Q3DIS6OTHa)))
firstup(names(table(Analysis_Domain_D03a$Q3DIS6OTHb)))
unique(firstup(c(names(table(Analysis_Domain_D03a$Q3DIS6OTHa)), names(table(Analysis_Domain_D03a$Q3DIS6OTHb)))))


### Average clinical dosage by intraoperative treatment method (in mL)
library(kableExtra)
papeR::summarize(Analysis_Domain_D03a %>% select(Q4ITM1,Q4ITM2, Q4ITM3, Q4ITM4),
                 type = "numeric", test = F, show.NAs = T) %>%
  kable(caption = "Descriptive statistics of continious variables", format = "html") %>%
  kable_styling(latex_options = "striped")
```


## Complications and Adverse Events Analysis


```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}

## cat(paste(colnames(read.delim("clipboard")), collapse=","), "\n")
  
### Select variables
Analysis_Domain_D03b <- Analysis_Domain_D02 %>%
  select(HOSCODE, Q1NCASE, 
         Q5RESIDUE,Q5YESCAT1,Q5YESCAT2,Q5YESCAT3,Q5YESCAT4,Q5YESCAT5,Q5YESCAT6,Q5YESCAT7OTH,Q5YESCAT7,
         Q6COMP,Q6COMPN,
         Q7HAZARD,Q7HAZNAME,Q7HAZN,
         Q8AE,Q8AEN) %>%
  mutate(Q5YES_N = Q5YESCAT1+Q5YESCAT2+Q5YESCAT3+Q5YESCAT4+Q5YESCAT5+Q5YESCAT6+Q5YESCAT7)

### Intraocular residue of Perfluorocarbon RT DECALIN after surgery
sum(Analysis_Domain_D03b$Q5YES_N)
sum(Analysis_Domain_D03b$Q5YES_N)/N_case  

### Intraocular residual location1 
apply(Analysis_Domain_D03b %>% select(Q5YESCAT1,Q5YESCAT2,Q5YESCAT3,Q5YESCAT4,Q5YESCAT5,Q5YESCAT6,Q5YESCAT7), 2, sum)
apply(Analysis_Domain_D03b %>% select(Q5YESCAT1,Q5YESCAT2,Q5YESCAT3,Q5YESCAT4,Q5YESCAT5,Q5YESCAT6,Q5YESCAT7), 2, sum) %>% prop.table() %>% round(3)
 
### Complications due to residual of Perfluorocarbon RT DECALIN1
table(Analysis_Domain_D03b$Q6COMP) %>% prop.table()  
 
### Other eye hazards caused by chemical or biological pollutants
table(Analysis_Domain_D03b$Q7HAZARD) %>% prop.table()  

### Adverse events or complications due to RT DECALIN residues
table(Analysis_Domain_D03b$Q8AE) %>% prop.table() 
sum(Analysis_Domain_D03b$Q8AEN)/N_case  
 
```

## Satisfaction Analysis

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}

## cat(paste(colnames(read.delim("clipboard")), collapse=","), "\n")
  
### Select variables
Analysis_Domain_D03c <- Analysis_Domain_D02 %>%
  select(HOSCODE, Q1NCASE, Q10EFFECT,Q11SUREXP,Q12SAFETY,Q13ECO,Q14RECOM) 

table(Analysis_Domain_D03c$Q10EFFECT) %>% prop.table()  
table(Analysis_Domain_D03c$Q11SUREXP) %>% prop.table()  
table(Analysis_Domain_D03c$Q12SAFETY) %>% prop.table()   
table(Analysis_Domain_D03c$Q13ECO) %>% prop.table()  
table(Analysis_Domain_D03c$Q14RECOM) %>% prop.table() 
```





# Grapic analysis

## Sex

<!-- ---------------------------------------------------------------------- -->
<!--          Example View missing values and descriptive statistics        -->
<!-- ---------------------------------------------------------------------- -->

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
data <- data.frame(
  group=c("Female", "Male"),
  value=c(6267,9315 )
)
# Compute the position of labels
data <- data %>% 
  arrange(desc(group)) %>%
  mutate(prop = paste(round(value / sum(data$value) *100,1),"%"),
         group = paste(group,"(",prop, ")")) 
### Converting all chr variables
data[sapply(data, is.character)] <- lapply(data[sapply(data, is.character)], as.factor)

# Basic piechart
ggplot(data, aes(x="", y=value, fill=group)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_fill_discrete(name = "Gender") +
  scale_color_brewer(palette="Dark2") +
  labs(title="Demographic: Gender",
       x="",
       y="") +
  theme(legend.position="bottom")+
  theme(axis.line = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid  = element_blank())
```



## Age
 

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
data <- data.frame(
  group=c("< 18 years old", "18 to 59 years old", "\u2265 60 years old"),
  value=c(803, 9154, 5625 )
)



# Compute the position of labels
data <- data %>% 
  arrange(desc(group)) %>%
  mutate(prop = paste(round(value / sum(data$value) *100,1),"%"),
         group = paste(group,"(",prop, ")")) 

 
 data$group

  


# Basic piechart
ggplot(data, aes(x="", y=value, fill=group)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_color_brewer(palette="Dark2") +
  scale_fill_discrete(name = "Age", breaks=c('< 18 years old ( 5.2 % )', '18 to 59 years old ( 58.7 % )', '\u2265 60 years old ( 36.1 % )'))+
  labs(title="Demographic: Age",
       x="",
       y="") +
  theme(legend.position="bottom")+
  theme(axis.line = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid  = element_blank())
```


## Disease 
 

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
data <- data.frame(
  group=c("Severe retinal detachments", 
          "Traumatic retinal detachments", 
          "Giant retinal tears",
          "Other retinal detachments",
          "Lifting of the subluxated lens",
          "Other diseases"),
  value=c(7468 , 2843 , 2889 ,827 ,1320 ,235 )
)
# Compute the position of labels
data <- data %>% 
  arrange(desc(group)) %>%
  mutate(prop = paste(round(value / sum(data$value) *100,1),"%"),
         group = paste(group,"(",prop, ")")) 
### Converting all chr variables
data[sapply(data, is.character)] <- lapply(data[sapply(data, is.character)], as.factor)

data$group
# Basic piechart
ggplot(data, aes(x="", y=value, fill=group)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_color_brewer(palette="Dark2") +
  scale_fill_discrete(name = "Disease",
                      breaks=c("Severe retinal detachments ( 47.9 % )",
                              "Traumatic retinal detachments ( 18.2 % )",
                              "Giant retinal tears ( 18.5 % )",
                              "Other retinal detachments ( 5.3 % )",
                              "Lifting of the subluxated lens ( 8.5 % )",
                              "Giant retinal tears ( 18.5 % )",
                              "Other diseases ( 1.5 % )")) +
  labs(title="Disease requiring aid treatment with Perfluorocarbon RT DECALIN",
       x="",
       y="") +
  theme(legend.position="right")+
  theme(axis.line = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid  = element_blank())
```


## Intraocular residual location

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
data <- data.frame(
  group=c("Anterior chamber", 
          "Vitreous cavity", 
          "Above the retina",
          "Under the retina",
          "Under the macular retina",
          "Under the posterior capsule of lens",
          "Other location"),
  value=c(16,35,39,21,4,11,0)
)
# Compute the position of labels
data <- data %>% 
  arrange(desc(group)) %>%
  mutate(prop = paste(round(value / sum(data$value) *100,1),"%"),
         group = paste(group,"(",prop, ")")) 
### Converting all chr variables
data[sapply(data, is.character)] <- lapply(data[sapply(data, is.character)], as.factor)

data$group

# Basic piechart
ggplot(data, aes(x="", y=value, fill=group)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_color_brewer(palette="Dark2") +
  scale_fill_discrete(name = "Intraocular residual location",
                      breaks=c("Anterior chamber ( 12.7 % )",
                              "Vitreous cavity ( 27.8 % )",
                              "Above the retina ( 31 % )",
                              "Under the retina ( 16.7 % )",
                              "Under the macular retina ( 3.2 % )",
                              "Under the posterior capsule of lens ( 8.7 % )",
                              "Other location ( 0 % )"),
                      labels=c("Anterior chamber ( 12.7 % )",
                              "Vitreous cavity ( 27.8 % )",
                              "Above the retina ( 31.0 % )",
                              "Under the retina ( 16.7 % )",
                              "Under the macular retina ( 3.2 % )",
                              "Under the posterior capsule of lens ( 8.7 % )",
                              "Other location ( 0.0 % )"))+
  labs(title="Intraocular residual location",
       x="",
       y="") +
  theme(legend.position="right")+
  theme(axis.line = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid  = element_blank())
```
 

## Satisfaction 

### 1. Overall clinical therapeutic effect evaluatio

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
data <- data.frame(
  group=c("Very effective", 
          "Effective", 
          "Uncertain",
          "Basically ineffective",
          "Completely ineffective"),
  value=c(19,12,0,0,0)
)
data$group <- factor(c("Very effective", 
          "Effective", 
          "Uncertain",
          "Basically ineffective",
          "Completely ineffective"),
          levels = c("Very effective", 
          "Effective", 
          "Uncertain",
          "Basically ineffective",
          "Completely ineffective"))

# Compute the position of labels
data <- data %>% 
  arrange(desc(group)) %>%
  mutate(prop = paste(round(value / sum(data$value) *100,1),"%"))

data$prop <- as.factor(data$prop)

data %>% ggplot(aes(x = group, y = value)) +
  geom_bar(stat = "identity", fill="#66a3ff", alpha=.6, width=.4) +
  geom_text(aes(label=prop, vjust=1.4,hjust=0.5)) +
  ylab("Number of events") +
  xlab("Overall clinical therapeutic effect evaluation") +
  theme_bw()
```


### 2. Satisfaction with the surgical experience

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
data <- data.frame(
  group=c("Very satisfied", 
          "Satisfied", 
          "Uncertain",
          "Dissatisfied",
          "Very dissatisfied"),
  value=c(16,15,0,0,0)
)

data$group <- factor(c("Very satisfied", 
          "Satisfied", 
          "Uncertain",
          "Dissatisfied",
          "Very dissatisfied"),
          levels = c("Very satisfied", 
          "Satisfied", 
          "Uncertain",
          "Dissatisfied",
          "Very dissatisfied"))


# Compute the position of labels
data <- data %>% 
  arrange(desc(group)) %>%
  mutate(prop = paste(round(value / sum(data$value) *100,1),"%"))

data$prop <- as.factor(data$prop)


data %>% ggplot(aes(x = group, y = value)) +
  geom_bar(stat = "identity", fill="#4d94ff", alpha=.6, width=.4) +
  geom_text(aes(label=prop, vjust=1.4,hjust=0.5)) +
  ylab("Number of events") +
  xlab("Satisfaction with the surgical experience") +
  theme_bw()
```



### 3. Satisfaction with the visual field clarity and surgical safety 

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
data <- data.frame(
  group=c("Very satisfied", 
          "Satisfied", 
          "Uncertain",
          "Dissatisfied",
          "Very dissatisfied"),
  value=c(16,15,0,0,0)
)
data$group <- factor(c("Very satisfied", 
          "Satisfied", 
          "Uncertain",
          "Dissatisfied",
          "Very dissatisfied"),
          levels = c("Very satisfied", 
          "Satisfied", 
          "Uncertain",
          "Dissatisfied",
          "Very dissatisfied"))

# Compute the position of labels
data <- data %>% 
  arrange(desc(group)) %>%
  mutate(prop = paste(round(value / sum(data$value) *100,1),"%"))

data$prop <- as.factor(data$prop)

data %>% ggplot(aes(x = group, y = value)) +
  geom_bar(stat = "identity", fill="#3385ff", alpha=.6, width=.4) +
  geom_text(aes(label=prop, vjust=1.4,hjust=0.5)) +
  ylab("Number of events") +
  xlab("Satisfaction with the visual field clarity and surgical safety ") +
  theme_bw()
```




### 4. Economic applicability and patient satisfaction with finical burden

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
data <- data.frame(
  group=c("Very satisfied", 
          "Satisfied", 
          "Uncertain",
          "Dissatisfied",
          "Very dissatisfied"),
  value=c(19,10,1,1,0)
)
data$group <- factor(c("Very satisfied", 
          "Satisfied", 
          "Uncertain",
          "Dissatisfied",
          "Very dissatisfied"),
          levels = c("Very satisfied", 
          "Satisfied", 
          "Uncertain",
          "Dissatisfied",
          "Very dissatisfied"))

# Compute the position of labels
data <- data %>% 
  arrange(desc(group)) %>%
  mutate(prop = paste(round(value / sum(data$value) *100,1),"%"))

data$prop <- as.factor(data$prop)

data %>% ggplot(aes(x = group, y = value)) +
  geom_bar(stat = "identity", fill="#005ce6", alpha=.6, width=.4) +
  geom_text(aes(label=prop, vjust=1.4,hjust=0.5)) +
  ylab("Number of events") +
  xlab("Economic applicability and patient satisfaction with finical burden") +
  theme_bw()
```





### 5. Recommendation using RT DECALIN

```{r,echo = F,message = FALSE, error = FALSE, warning = FALSE}
data <- data.frame(
  group=c("Definitely", 
          "Probably", 
          "Uncertain",
          "Unwilling",
          "Definitely not"),
  value=c(23,8,0,0,0)
)
data$group <- factor(c("Definitely", 
          "Probably", 
          "Uncertain",
          "Unwilling",
          "Definitely not"),
          levels = c("Definitely", 
          "Probably", 
          "Uncertain",
          "Unwilling",
          "Definitely not"))

# Compute the position of labels
data <- data %>% 
  arrange(desc(group)) %>%
  mutate(prop = paste(round(value / sum(data$value) *100,1),"%"))

data$prop <- as.factor(data$prop)

data %>% ggplot(aes(x = group, y = value)) +
  geom_bar(stat = "identity", fill="#0047b3", alpha=.6, width=.4) +
  geom_text(aes(label=prop, vjust=1.4,hjust=0.5)) +
  ylab("Number of events") +
  xlab("Recommendation using RT DECALIN") +
  theme_bw()
```
