---
title: 'R Shiny.Note'
author: "Zehui Bai"
date: '`r format(Sys.time())`'
output:
  html_document:
    df_print: paged
    number_sections: no
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
  pdf_document:
    toc: yes
fontsize: 10pt
editor_options:
  chunk_output_type: console
colorlinks: yes
---

```{r setup, include=FALSE, echo = FALSE,message = FALSE, error = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# <!-- ---------------------------------------------------------------------- -->
# <!--                    1. load the required packages                       -->
# <!-- ---------------------------------------------------------------------- --> 

## if(!require(psych)){install.packages("psych")}

packages<-c("tidyverse", "knitr", "papeR")
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}
ipak(packages)
# <!-- ---------------------------------------------------------------------- --> 


# <!-- ---------------------------------------------------------------------- -->
# <!--                        2. Basic system settings                        -->
# <!-- ---------------------------------------------------------------------- -->
setwd(dirname(rstudioapi::getSourceEditorContext()$path))
getwd()
Sys.setlocale("LC_ALL","English")

### get the path
# rstudioapi::getSourceEditorContext()$path
# dirname(rstudioapi::getSourceEditorContext()$path)

### set working directory
# getwd()
# setwd("c:/Users/zbai/Desktop")
# Sys.setlocale("LC_ALL","English")

### get the R Version
# paste(R.Version()[c("major", "minor")], collapse = ".")

### convert backslash to forward slash 
# scan("clipboard",what="string")
# gsub('"', "", gsub("\\\\", "/", readClipboard()))
# <!-- ---------------------------------------------------------------------- --> 



# <!-- ---------------------------------------------------------------------- -->
# <!--     3. Load the SASmarkdown package if the SAS output is required      -->
# <!-- ---------------------------------------------------------------------- -->
# library(SASmarkdown)
# ### Set SAS output
# ### Reset engine to R
# saspath <- "C:/SASHome/SASFoundation/9.4/sas.exe"
# sasopts <- "-nosplash -linesize 75"
# knitr::opts_chunk$set(engine="sashtml", engine.path=saspath,
#         engine.opts=sasopts, comment=NA)
# 
# # run these commands to convince yourself that
# # within this knitr session the engine changed.
# knitr::opts_chunk$get()$engine
# knitr::opts_chunk$get()$engine.path
# knitr::opts_chunk$get()$engine.opts
# <!-- ---------------------------------------------------------------------- -->



# <!-- ---------------------------------------------------------------------- -->
# <!--                         4. Import the datasets                         -->
# <!-- ---------------------------------------------------------------------- -->
### Import csv data
# pfad <- "~/Desktop/SASUniversityEdition/myfolders/Daten"
# mydata1 <- read.csv(file.path(pfad, "yourcsv_data.csv"), 
#                     sep=";", 
#                     header=TRUE)   

### Import xlsx data
# library(readxl)
# mydata2 <- read_excel("C:/Users/zbai/Documents/GitHub/R-Projects/SAS/Yimeng/results-text.xlsx")

### Import sas data
# library(sas7bdat)
# mydata3 <- read.sas7bdat("~/Desktop/SASUniversityEdition/myfolders/Daten/uis.sas7bdat")

### Import from copyboard
# copdat <- read.delim("clipboard")
# Data_D01 <- copdat

# <!-- ---------------------------------------------------------------------- -->
# <!--                           5. Some Tools                                -->
# <!-- ---------------------------------------------------------------------- -->

## To check out vignettes for one specific package
# browseVignettes("ggplot2")


# <!-- ---------------------------------------------------------------------- -->
```



```{r mind map,echo = F,message = FALSE, error = FALSE, warning = FALSE}
## Convert to mind map text, markdown outline, R script, and HTML widget ####
library(mindr)
# text -> widget
# input <- c("# Chapter 1", "## Section 1.1", "### Section 1.1.1", "## Section 1.2", "# Chapter 2")
# mm(from = input, root = "mindr")


input <- rstudioapi::getSourceEditorContext()$path
mm(from = input, type = 'file', widget_name = 'mindrtest.html')

## file.show(input) # Open the input file with the default program, if any
## Old
# input_txt <- readLines(input, encoding = "UTF-8")
## Convert to mind map text, markdown outline, R script, and HTML widget ####
# mm_output <- mm(input_txt, 
#                 output_type = c("widget"),
#                 root = "")
# mm_output$widget
```

# ui

## Shinydashboard

* dashboardHeader: Create a header for a dashboard page
* dashboardSidebar: Create a dashboard sidebar
* dashboardBody: The main body of a dashboard page
  + tabItems: A container for tab items
    + The familiar sidebarLayout() described above makes use of Shiny’s lower-level grid layout functions. Rows are created by the fluidRow() function and include columns defined by the column() function. Column widths are based on the Bootstrap 12-wide grid system, so should add up to 12 within a fluidRow() container.
    + sidebarLayout: sidebarLayout(sidebarPanel, mainPanel, position = c("left", "right")), Create a layout (sidebarLayout()) with a sidebar (sidebarPanel()) and main area (mainPanel()).
    + sidebarPanel，The sidebarPanel() containing input controls.
    + varSelectInput: Create a select list that can be used to choose a single or multiple items from the column names of a data frame. 'varSelectInput( inputId, label, data, selected = " ")'
    + numericInput: Create an input control for entry of numeric values. 'numericInput( inputId, label, value)'
    + radioButtons: Create a set of radio buttons used to select an item from a list. 'radioButtons(inputId, label, choices, selected = NULL)'
    + mainPanel: 	Create a main panel containing output elements that can in turn be passed to sidebarLayout.
    + tabsetPanel: Tabsets are useful for dividing output into multiple independently viewable sections.
    




# sever

